---
title: "US College data"
subtitle: "A Comprehensive Overview of Various Metrics"
author: "; Naga Kartheek Peddisetty"
format: dashboard

theme:
  - cosmo
  - custom.scss
  - quarto.scss

execute: 
  echo: false
---

```{r setup, include=FALSE}

# Load necessary libraries
library(ISLR2)
library(tidyverse)
```

```{r}
# Load the College dataset
data(College)
college <- College
college <- college %>%
  mutate(Private = ifelse(Private == "No", "Public", "Private"))
college$Grad.Rate[college$Grad.Rate > 100] <- 100
```

# Statistics
## Row 

::: {.valuebox icon="star" color="vb-info"}
Private Universities

`r sum(College$Private=="Yes")`
:::

::: {.valuebox icon="plus-circle" color="vb-warning"}
Public Universities

`r sum(College$Private=="No")`
:::

::: {.valuebox icon="star" color="vb-danger"}
New students from top 10% of H.S class

`r format(sum(College$Top10perc), big.mark = ",")`
:::

::: {.valuebox icon="plus-circle" color="vb-success"}
Overall Grad.Rate

`r round(mean(College$Grad.Rate),2)`
:::


## Row
### Column {.tabset width="50%"}
```{r}
#| title: Box plot
# Create a box plot for Graduation Rate (Grad.Rate) by Private/Public universities
ggplot(college, aes(x = Private, y = Grad.Rate, fill = Private)) +
  geom_boxplot() +
  labs(title = "Box Plot of Graduation Rate by University Type",
       x = "University Type",
       y = "Graduation Rate (%)",
       fill = "University Type") +
  theme_minimal()
```

```{r}
#| title: Scatter plot Analysis
# Create a scatter plot for Graduation Rate vs. Out-of-State Tuition
ggplot(college, aes(x = Outstate, y = Grad.Rate, color = Private)) +
  geom_point() +
  facet_wrap(~Private, ncol = 2) +
  labs(x = "Out-of-State Tuition ($)", y = "Graduation Rate(%)", color = "Private University") +
  ggtitle("Graduation Rate vs. Out-of-State Tuition")
```

### Column {.tabset width="50%"}
```{r}
#| title: Histogram of Grad.Rate
ggplot(college, aes(x = Grad.Rate)) +
  geom_histogram(binwidth = 5, fill = "salmon", color = "black") +
  labs(title = "Histogram of Graduation Rate",
       x = "Graduation Rate (%)",
       y = "No. of Universities") +
  theme_minimal()
```

```{r}
#| title: Line Plot
# Line plot with smoother for Percent Faculty with Terminal Degree vs. Percent Faculty with PhD
ggplot(college, aes(x = Terminal, y = PhD, color = Private)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ Private, ncol = 2) +
  labs(title = "PhD vs. Terminal Degree Percentage",
       x = "Percent Faculty with Terminal Degree",
       y = "Percent Faculty with PhD",
       color = "Private University") +
  theme_minimal()
```

# About Dataset

::: card
::: card-header
What is the data about ??
:::
::: card-body
US College data set based on US news and World report.

**Dimensions of the data set:**

• This data set consists of 777 observations and 18 variables. Primarily used to demonstrate statistical and machine learning methods.

**Important Attributes description:**

> Private: Categorical variable indicating whether the college is private ("Yes") or public ("No").
>
> Apps: Number of applications received (integer).
>
> Accept: Number of applicants accepted (integer).
>
> Enroll: Number of new students enrolled (integer).
>
> Outstate: Out-of-state tuition fees (USD).
>
> F.Undergrad: The number of full-time undergraduates (integer).
>
> P.Undergrad: The number of part-time undergraduates (integer). 
>
> Room.Board: Room and board costs (USD).
>
> Books: Estimated book costs (USD).
>
> Grad.Rate: Graduation rate (percentage).
:::
:::
